---
title: "R Exercises"
author: "Sofi Khachatryan"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
```{r}
library(stringr)
library(dplyr)
library(ggplot2)
```

**Create a function, which will get an input a text character vector, where there should be phone numbers. The function should extract all numbers that are valid Armenian numbers (check that it starts with 0 and does not have letters in it and number of digits is 9: that is enough), bring all valid numbers to one format (either with "-"s or without, either with spaces or without)**

```{r}
phone_numbers <- c("091-234931", "5214698", "a033255687", "077-45-89-96", "256347077", "047155266", "05523365d", "096900611", "096-900611", "11008005", "077 33 70 91")

get_formatted_numbers <- function(txts) {
  # First subset not to have any numbers with letters
  txts <- txts[!str_detect(txts, "[a-zA-z]")]
  # Delete all -s and spaces
  txts <- str_replace_all(txts, "-|\\s+", "")
  # subset by length and starting with 0 
  txts <- txts[nchar(txts) == 9 & str_sub(txts, start=1, end=1)=="0"]
  
  return (txts)
}

formatted_phone_numbers <- get_formatted_numbers(phone_numbers)
formatted_phone_numbers
```

**Having the result of previous exercise, count how many phone numbers belong to Beeline company, if the codes for Beeline company are "091", "099", "047", "043", "096".**      

```{r}
beeline_codes <- c("091", "099", "047", "043", "096")

pattern <- paste0("^", beeline_codes)
pattern <- paste0(pattern, collapse="|")

beeline_count <- sum(str_count(formatted_phone_numbers, pattern=pattern))
beeline_count
```



You are given a video games dataset containing information about popular video games, their sales in North America, Europe, Japan and globally in the world. 

**Import the *video_games_2.csv* dataset into R. Print the structure of the dataset, by 1-2 sentences write what you see in the structure.**
```{r}
video_games_df <- read.csv("video_games_2.csv")
str(video_games_df)
```

As we can see the dataset consists of 7586 observations and 15 features. There are 7 columns that are characters initially and 8 features that are numeric.


**Using dplyr package subset the dataframe by following these instructions:  
- remove columns Publisher, JP_Sales (Sales in Japan), Critic_Count, User_Count and Developer,  
- multiply the numbers in NA_Sales, EU_Sales and Global_Sales by 1 million as they are given in millions of sales,  
- include only those observations for which NA_Sales>=20000, EU_Sales>=20000 and Rating is among "E", "M", "T", "E10+", "AO" (the meanings of abbrevations: Everyone, Mature, Teen, Everyone 10+, Adults Only, they are used in the dataframe by the abbrevations)** 


```{r}
video_games <- video_games_df %>%
  select(-c(Publisher, JP_Sales, Critic_Count, User_Count, Developer)) %>%
  mutate(NA_Sales = NA_Sales * 1e+6,
         EU_Sales = EU_Sales * 1e+6, 
         Global_Sales = Global_Sales * 1e+6) %>%
  filter(Rating %in% c("E", "M", "T", "E10+", "AO") & NA_Sales>=20000 & EU_Sales>=20000)
str(video_games)
```
  

**Imagine that there are validation rules for 2 columns of our dataframe: "Critic_Score"s can be integer values from 0 to 100 and "User_Score"s can be integers from 1 to 10.** 
```{r} 
video_games_valid <- video_games %>%
  mutate(User_Score = round(User_Score), 
         Critic_Score = round(Critic_Score)) %>%
  filter(User_Score %in% c(1:10), 
         Critic_Score %in% c(0:100))
```

**Use subsetted data *video_games_valid* and create a boxplot displaying how *User_Score*s and *Critic_Score*s are interconnected. Use *User_Score* for the x-axis and as it is a numerical type but has discrete value range (1, 2, ..., 10), convert it to Factor datatype, then create the boxplot to have better results. Make the boxplot's edges *red* and boxplot's background *tomato*. **

```{r}
ggplot(data=video_games_valid, aes(x=as.factor(User_Score), y=Critic_Score)) +
  geom_boxplot(fill="tomato", col="red") +
  labs(title="Interconnection between User scores and Critics score", 
       x="User Score", 
       y="Critic Score")
```
It can be observed from the graph that as User Score increases, Critic Score tends to increase as well.

```{r}
cor(video_games_valid$User_Score, video_games_valid$Critic_Score)
```
As we can see from both the graph and correlation result,the correlation between two features is positive as one might suspect.


**Construct a scatterplot showing how the EU_Sales (Europe Sales) of the game is dependent of NA_Sales (North American Sales). Use subsetted dataframe video_games_valid** 


```{r}
options(scipen=999)
ggplot(data=video_games_valid, aes(x=NA_Sales, y=EU_Sales)) + 
  geom_point(alpha = 0.2, color='black')+
  labs(title="North American Sales dependence from the Europe Sales", x="North American Sales", y="Europe Sales")+
  coord_cartesian(ylim=c(0, 1e+7), xlim=c(1, 1e+7)) +
  theme_minimal()
```
One can see from the graph, that the correlation between two variables is mainly positive, but there are also cases, when the video game had much more sales in North America than in Europe, and the reverse.



**Create a barplot to find how many games of each Genre there are in the dataset. Use either *valid_video_games* or your initial dataset. State what are the top 3 Genres according to the graph. ** 

```{r}
ggplot(video_games_valid, aes(x=Genre)) +
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Distribution of the games by Genre", x="Genre Names", y="Number of games")
```
From the histogram, its easy to see that top 3 genres are Action, Sports and Shooter.

The exercise to try:
```{r}
video_games_valid %>%
  group_by(Genre) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(Genre, -count), y = count)) + geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Distribution of the games by Genre ordered by count", x="Genre Names", y="Number of games")
```

*show the number of video games in each genre in descending order according to the dataset.*
```{r}
vg_genres <- video_games_valid %>%
  group_by(Genre) %>%
  summarise(count=n()) %>%
  arrange(desc(count)) 
vg_genres
```

Define the *Rating* column as Factor (if it is not already so) and use faceting (facet_grid) to plot scatterplot *User_Score* vs *NA_Sales* (North America Sales) for different Ratings. Zoom the graph so User_Score is between 5 and 10, and North American Sales are from 0 to 15500000. Do not forget about the titles and interpretations. (7 points)

```{r}
video_games$Rating <- factor(video_games$Rating)
ggplot(video_games, aes(x=User_Score, y=NA_Sales)) +
  geom_point() + facet_grid(.~Rating) +
  labs(title="User score of the game VS the North America Sales for different Ratings", 
       x="Score given by the User", y="North America Sales")+
  coord_cartesian(xlim=c(5, 10), ylim=c(0, 1.55e+7))
```
User Scores and North America Sales variables are dependent and as one increases the other tends to increase as well. What is more, we can observe the differences between Ratings as well. For instance, we can see that games with rating 'Everyone' has games with the highest sales in North America.


Make these design changes on the plot from the previous exercise:   
- x axis name -- "Score given by the User" color: "khaki3", bold, size: 15,  
- y axis name -- "North America Sales" color: "khaki3", bold, size: 15,  
- points (shape - square, color: "brown", size: 1.5),  
- title of the plot -- "User score of the game VS the North America Sales for different Ratings",  
- panel background color: "beige",  
- axis texts bold black.  

```{r}
ggplot(video_games, aes(x=User_Score, y=NA_Sales))+
  geom_point(size = 1.5, col="brown", shape = 0)+facet_grid(.~Rating) +
  labs(title="User score of the game VS the North America Sales for different Ratings", 
       x="Score given by the User", y="North America Sales")+
  coord_cartesian(xlim=c(5, 10), ylim=c(0, 1.55e+7)) +
  
  theme(axis.text.x = element_text(color = "black", face="bold"),
      axis.text.y = element_text(color = "black", face="bold"),
      axis.title.x = element_text(size = 15, color = "khaki3", face="bold"),
      axis.title.y = element_text(size = 15, color = "khaki3", face="bold"),
      panel.background = element_rect(fill = 'beige')) 
```


 We are interested in the number of video games developed for platforms *'PS2', 'X360', 'PS3'* for different years.  
- Make other platforms, that are not from these 3 as "Other" using dplyr (Hint: ifelse statement).  
- Remove all observations from dataframe which have any NA values (Hint: ?complete.cases).  
- Use faceting to draw the distribution of games for each year for each platform. Make text on "x" axis vertical and size=5.   


```{r}
top_platforms <- c('PS2', 'X360', 'PS3')
video_games %>% 
  mutate(Platform = ifelse(Platform %in% top_platforms, Platform, 'Other')) %>%
  filter(complete.cases(.) ) %>%
  group_by(Platform, Year) %>%
  summarise(Count=n()) %>%
  ggplot(aes(x=Year, y=Count))+geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90, size = 5)) +
  labs(title="The distribution of games for each year for each platform", 
       x="Years", 
       y="Number of games") +
  facet_grid(.~Platform)
```
In all platforms we can see that there is a peak in the middle. The number of games start to increase, reach their maximum point and tend to decrease.
PS2 platform reachs its peak at 2005s when it has almost 180 games.
PS3 platform reachs its peak at 2010s when it has almost 110 games.
X360 platform reachs its peak at 2010s when it has almost 110 games.
And other platforms reached the pick at 2005s when they had almost 270 games.

*countries.csv* dataset to R by reading it. Then print first 4 observations in the dataset. Next state how many observations and how many features the dataframe has? Are all features numeric ones? 

```{r}
countries <- read.csv("countries.csv", stringsAsFactors = F)
head(countries, n=4)
```

For telling how many observations and features, the dataframe has, we can use the *str* function.
```{r}
str(countries)
```

We can see that the dataframe has 227 observations and 20 variables. We also can see that almost all features are numeric except of the Country and Region.


The column Population shows the population for all the countries. Find the maximum and minimum population values in the dataset. 

```{r}
min(countries$Population)
```

```{r}
max(countries$Population)
```


Having the min and max values, we can subset the dataframe accordingly, and choose the Country column, to see names of desired countries. 
```{r}
min_pop_country <- countries[countries$Population == 7026, "Country"]
min_pop_country
```

```{r}
max_pop_country <- countries[countries$Population == 1313973713, "Country"]
max_pop_country
```

As one would expect, the most populated country is China, and the one with smallest population is St Pierre & Miquelon.


Now suppose we want to consider only those countries, which are in the region *C.W. OF IND. STATES*. Subset the dataframe as follows. Name the new dataframe *CIS_countries*.

```{r}
CIS_countries <- countries[countries$Region == "C.W. OF IND. STATES", ]
str(CIS_countries)
```

And we can see that there are 12 countries that are CIS members according to the dataset. 


Consider the climate of the CIS countries. Calculate the mean and standard deviation of the feature. What can you say about the results? 

```{r}
mean_cis_climate <- mean(CIS_countries$Climate, na.rm = TRUE)
mean_cis_climate
```

```{r}
std_cis_climate <- sd(CIS_countries$Climate, na.rm = TRUE)
std_cis_climate
```

We can see that standard deviation is quite big considering the fact that the range of Climate features is from 1 to 4, so the climate of the CIS countries are quite different.  

The difference between the birth rate and the death rate of a country or place is called the natural increase. The natural increase is calculated by subtracting the death rate from the birth rate. *Natural increase = birth rate - death rate.* Calculate Natural Increase and keep it in the dataframe under a column *NaturalIncrease*. Then find which countries have the highest and lowest natural increase.
```{r}
countries$NaturalIncrease <- countries$Birthrate - countries$Deathrate
min_NI <- min(countries$NaturalIncrease, na.rm = T)
max_NI <- max(countries$NaturalIncrease, na.rm = T)
min_NI
max_NI
```
```{r}
countries_NI <- countries[!is.na(countries$NaturalIncrease),]
min_NI_country <- countries_NI[countries_NI$NaturalIncrease == min_NI, "Country"]
max_NI_country <- countries_NI[countries_NI$NaturalIncrease == max_NI, "Country"]
min_NI_country
max_NI_country
```

As we can see they are not the same countries with min and max population. Bostwana has the least Natural Increase, and Gaza Strip has the biggest value for Natural Increase. 


*Plot a scatterplot using between Population and Area_sqm*. 

```{r}
library(ggplot2)
ggplot(data=countries, aes(x=Population, y=Area_sqm)) +
  geom_point() + 
  labs(title = "Scatterplot between Population and Are of countries", 
       x = "Population of the country", y = "Area of the country") +
  theme_minimal() 
```
As we can see logically countries with extremely high population have bigger area. Also one can see, that there are 6 countries that have larger area than most of the countries.  

Because of the outliers, we can interpret the graph better if we zoom in:

```{r}
ggplot(data=countries, aes(x=Population, y=Area_sqm)) +
  geom_point() + 
  labs(title = "Scatterplot between Population and Are of countries", 
       x = "Population of the country", y = "Area of the country") +
  theme_minimal() +
  xlim(0, 150000000) + ylim(0, 3000000)
```

In zoomed-in version of the graph, more points can be seen clearly, and we can see that there are both countries that have bigger area and bigger population, and also countries that have high population but small area, and the inverse. 





